---
title: "UPDRSIII.sobgroups_progression"
author: "AMCalejandro"
date: "23/7/2021"
output: html_document
---


```{r setup, include=FALSE}
library(tidyverse)      # For tidy manipulation of data
library(here)           # For file path construction
library(DT)             # To display data tables
library(readxl)         # To read excel files
theme_rhr <-  theme_bw(base_family = "Helvetica") + 
  theme(panel.grid.major.x = element_blank(),
        legend.position = "right",
        strip.text = element_text(size = 8),
        strip.text.y = element_text(angle = 90),
        axis.text.x = element_text(size = 8, angle = 90, hjust = 1, vjust = 0.5),
        axis.text.y = element_text(size = 8),
        axis.title.y = element_text(vjust = 0.6),
        axis.title = element_text(size = 10),
        panel.spacing = unit(0.1, "lines"))
knitr::opts_chunk$set(echo = T, warning = F, message= F)
```

> Aim: We want to study the differences in the progression of the MDS-UPDRS III derived outcomes ( total, limb and axial)


# Description

In order to see the rate of progression of the different PD motor symptoms, we are going to plot the mean MDS-UPDRS III derived outcomes for PPMI, PROBAND and OPDC across different time points. 
To do so, we are going to use the limb, axial, and total outcome variables that were scaled up on the workflow 2.MissingnessExploration.Imputation_part1.

# Ploting the differences in progression

Look at the difference in progression for the MDS-UPDRSIII derived measures 

# Data load

```{r}
# Three visits only
TPD_PDonly_3visits_long_updrsIII_imputed <- read.csv("../Data_processed/clinicalData_Imputed/TPD_0.8.imputed.txt")
OPDC_3visits_long_updrsIII_imputed <- read.csv("../Data_processed/clinicalData_Imputed/OPDC_0.8.imputed.txt")
PPMI_long_updrsIII_imputed <- read.csv("../Data_processed/clinicalData_Imputed/PPMI_0.8.imputed.txt")
DIGPD_long_updrsIII_imputed <- read.csv("../Data_processed/clinicalData_Imputed/DIGPD_0.8.imputed.txt")
PDSTAT_long_updrsIII_imputed <- read.csv("../Data_processed/clinicalData_Imputed/PDSTAT_0.8.imputed.txt")
PDBP = readRDS("../Data_processed/Rfiles_modelFitting.automationOutline/PDBP/PDBP_final.rds")[1:8]

# All visits up to 90 months
TPD_PDonly_AllVisits_long_updrsIII_imputed <- read.csv("../Data_processed/clinicalData_Imputed/TPD_0.8.imputed.AllVisits.txt")
OPDC_AllVisits_updrsIII_imputed <- read.csv("../Data_processed/clinicalData_Imputed/OPDC_0.8.imputed.AllVisits.txt")
PPMI_AllVisits_long_updrsIII_imputed <- read.csv("../Data_processed/clinicalData_Imputed/PPMI_0.8.imputed.AllVisits.txt")
DIGPD_AllVisits_long_updrsIII_imputed <- read.csv("../Data_processed/clinicalData_Imputed/DIGPD_0.8.imputed.AllVisits.txt")


```

# Data processing

```{r}
TPD_PDonly_3visits_long_updrsIII_imputed <- TPD_PDonly_3visits_long_updrsIII_imputed %>%
  dplyr::mutate(visit_number = recode(visit_number, `1` = "0", `4` = "1.5", `7` = "3")) %>%
  dplyr::rename(visit = visit_number) %>%
  dplyr::mutate(DF = "TPD")

OPDC_3visits_long_updrsIII_imputed <- OPDC_3visits_long_updrsIII_imputed %>%
  dplyr::mutate(visit = recode(visit, `1` = "0", `2` = "1.5", `3` ="3")) %>%
  dplyr::rename(ID = subjid) %>%
  dplyr::mutate(DF = "OPDC") %>%
  as_tibble()

PPMI_long_updrsIII_imputed <- PPMI_long_updrsIII_imputed %>%
  dplyr::filter(EVENT_ID %in% c("BL","V05","V08")) %>%
  dplyr::mutate(EVENT_ID = (recode(EVENT_ID, "BL" = "0", "V05" = "1.5", "V08" = "3"))) %>%
  dplyr::rename(c(visit = EVENT_ID, ID = PATNO)) %>%
  dplyr::mutate(DF = "PPMI")

DIGPD_long_updrsIII_imputed <- DIGPD_long_updrsIII_imputed %>%
  dplyr::mutate(visite = (recode(visite, `1` = "0", `2` = "1", `3` = "2", `4`="3"))) %>%
  dplyr::rename(ID = PATIENT) %>%
  dplyr::rename(visit = visite) %>%
  dplyr::mutate(DF = "DIGPD")

PDSTAT_long_updrsIII_imputed <- PDSTAT_long_updrsIII_imputed %>%
  dplyr::mutate(visit_number = recode(visit_number, `0` = "0", `12` = "1", `24` = "2" ))  %>%
  dplyr::rename(ID = StudyNumber) %>%
  dplyr::rename(visit = visit_number) %>%
  dplyr::mutate(DF = "PDSTAT")

PDBP <- PDBP %>%
  dplyr::filter(visit_month %in% c(0,12,18, 24,36)) %>%
  dplyr::mutate(visit_month = recode(visit_month, `0` = "0", `12` = "1", `18` = "1.5", `24` = "2", `36` = "3" ))  %>%
  dplyr::rename(visit = visit_month) %>%
  dplyr::mutate(DF = "PDBP") %>%
  group_by(ID) %>% filter(!n() == 1) %>% ungroup()

merged <- rbind(TPD_PDonly_3visits_long_updrsIII_imputed, OPDC_3visits_long_updrsIII_imputed,
                PPMI_long_updrsIII_imputed, DIGPD_long_updrsIII_imputed,
                PDSTAT_long_updrsIII_imputed, PDBP)
```


# Get the motor scores trayectories across cohorts


```{r}
my_mean = function(x) { mean(x, na.rm = T)}
my_sd = function(x) {sd(x, na.rm = T)}
meanAcrossVisits <- merged %>%
  group_by(visit, DF) %>%
  dplyr::summarise(across(ends_with("imputed"), list(mean = my_mean, sd = my_sd), .names = "{.col}.{.fn}")) %>%
  ungroup() %>%
  gather(UPDRS_type, score, -visit, -DF) %>%
  extract(UPDRS_type, c("UPDRS_type", "Metric"), "(.*)\\.(.*)") %>%
  spread(Metric, score) %>%
  mutate(UPDRS_type = 
           stringi::stri_replace_all_regex(UPDRS_type,
                                           pattern = c("UPDRSIIItotal_imputed", "UPDRSIIIlimb_imputed", "UPDRSIIIaxial_imputed"),
                                           replacement = c("Total", "Limb", "Axial"), vectorize = FALSE),
         DF = factor(DF, levels = c("TPD", "OPDC", "PPMI", "DIGPD", "PDSTAT", "PDBP")),
         visit = factor(visit, levels = c("0", "1","1.5", "2", "3")))

a = ggplot(meanAcrossVisits, aes(x = as.numeric(as.character(visit)), y = mean, color = UPDRS_type)) +
  geom_point(size = 5 ) +
  geom_line(size = 1.1) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd),
                size = 0.5, width = .1)  +
  facet_wrap(vars(DF), scales = "free") +

  
  theme_bw(base_size = 10) +
  labs(x = "Years", y = "MDS-UPDRS III Scores") +
  ggtitle("Average change in MDS-UPDRS III motor scores ") +
  theme(
    text = element_text(size = 35),
    plot.title = element_text(hjust = 0.5, size = 40, face = "bold", margin = margin(b = 20)),
    legend.title = element_text(size = 25),
    axis.title.x = element_text(margin = margin(t = 20), size = 40), 
    axis.title.y = element_text(margin = margin(r = 20), size = 40),
    strip.text = element_text(size=30))


png(file = "/home/amcalejandro/Kronos_downloads/motorscores_trayextpory.png", width=15,height=15,units="in",res=800)
a
dev.off()
```


# Furhter exploration up to 90 months from BL

Visualise the trayectories for longer period in TPD and OPDC

```{r}
# We modify the visit values for TPD to match with the OPDC df
TPD_PDonly_AllVisits_long_updrsIII_imputed <- TPD_PDonly_AllVisits_long_updrsIII_imputed %>% 
  dplyr::mutate(visit_number = recode(visit_number, `4` = 2L, `7` = 3L, `9` = 4L, `10` = 5L, `11` = 6L)) %>%
  dplyr::rename(visit = visit_number) %>%
  dplyr::mutate(DF = "TPD")

OPDC_AllVisits_updrsIII_imputed <- OPDC_AllVisits_updrsIII_imputed %>%
  dplyr::rename(ID = subjid) %>%
  dplyr::mutate(DF = "OPDC") %>%
  as_tibble()


merged_OPDC.TPD <- rbind(TPD_PDonly_AllVisits_long_updrsIII_imputed, OPDC_AllVisits_updrsIII_imputed)


TPD_Scaledmeans <- merged_OPDC.TPD %>%
  group_by(visit) %>%
  mutate_at(c("UPDRSIIItotal_imputed", "UPDRSIIIaxial_imputed", "UPDRSIIIlimb_imputed" ), ~scale(., center = FALSE)) %>%
  summarise(across(ends_with("imputed"), ~ mean(.x, na.rm = TRUE))) %>%
  ungroup() %>%
  gather(UPDRS_type, score, -visit)


ggplot(TPD_Scaledmeans, aes(x = visit, y = score, group = UPDRS_type, color = UPDRS_type)) +
  geom_point(size = 5 ) +
  geom_line(size = 1.1) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd),
                size = 0.5, width = .1)  +
  facet_wrap(vars(DF), scales = "free") +
  
  theme_bw(base_size = 10) +
  labs(x = "Years", y = "MDS-UPDRS III Scores") +
  ggtitle("Average change in MDS-UPDRS III motor scores ") +
  theme(
    text = element_text(size = 35),
    plot.title = element_text(hjust = 0.5, size = 40, face = "bold", margin = margin(b = 20)),
    legend.title = element_text(size = 25),
    axis.title.x = element_text(margin = margin(t = 20), size = 40), 
    axis.title.y = element_text(margin = margin(r = 20), size = 40),
    strip.text = element_text(size=30))
```

