---
title: 'Imputation of splitted UPDRS measures'
author: 
- name: "Alejandro M.Carrasco"
  affiliation: UCL
output: 
  html_document:
    css: ./font/missingexploration_fontsize.css
    code_folding: hide
    theme: paper
    highlight: kate
    df_print: paged
    fontsize: 20pt
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
library(tidyverse)      # For tidy manipulation of data
library(here)           # For file path construction
library(DT)             # To display data tables
library(readxl)         # To read excel files
library(VIM)            # For missing data exploration
library(kableExtra)     # For fancy data frame display
library(ggpubr)         # To display ggboxplots
library(naniar)         # reaplce values with NAs on dfs
theme_rhr <-  theme_bw(base_family = "Helvetica") + 
  theme(panel.grid.major.x = element_blank(),
        legend.position = "right",
        strip.text = element_text(size = 8),
        strip.text.y = element_text(angle = 90),
        axis.text.x = element_text(size = 8, angle = 90, hjust = 1, vjust = 0.5),
        axis.text.y = element_text(size = 8),
        axis.title.y = element_text(vjust = 0.6),
        axis.title = element_text(size = 10),
        panel.spacing = unit(0.1, "lines"))
knitr::opts_chunk$set(echo = T, warning = F, message= F)

```



# Description

At this stage we want to retrieve as many measures as possible for each time point. We will use the same strategy from the methods described in here: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5072275/ 
Sometimes, when patients miss even a single measure, the UPDRS_III total score is reported as missing. We will get less stringent and we will consider NAs measures only when patients miss more than 80% of the total terms over each MDS-UPDRS III derived outcomes (MDS-UPDRS III total, limb, and axial).
We will explore missingness of our outcome variable in order to understand whether data is MAR, MCAR or MNAR, and whether we can impute our data further with any software such as MICE.



# Data load

We load on out workspace all the data we are going to process

```{r}
Discovery_FU <- (read.csv("~/Data/WorkingDirectory/OPDC/clinical_data/Discovery_FU_data_with_onset_2021_04.csv"))
P3_TPD <- as_tibble(read.csv("~/Data/WorkingDirectory/PROBAND/clinical_data/P3_TPD_Version2_temp_06_16_2020.csv"))
PPMI_UPDRS<- as_tibble(read.csv("/home/amcalejandro/Data/WorkingDirectory/PPMI/clinical_data/MDS_UPDRS_Part_III.csv"))
PPMI_PD_IDs <- read.delim("/home/amcalejandro/Data/WorkingDirectory/PPMI/clinical_data/PD_IDs.txt", header=FALSE) 
PPMI_Original_Cohort <- read.csv("/home/amcalejandro/Data/WorkingDirectory/PPMI/clinical_data/PPMI_Original_Cohort_BL_to_Year_5_Dataset_Apr2020.csv")
PDSTAT_UPDRS <-  readxl::read_xlsx("/home/amcalejandro/Data/WorkingDirectory/PDSTAT/clinical_data/FUP_Export_CRF_MDSUPDRS_PartIII_Off.xlsx")
DIGPD <- readxl::read_xlsx("/home/amcalejandro/Data/WorkingDirectory/DIGPD_DATA/dig_pd_2019_11_Alejandro.xlsx")

```


# Description

In this section, for each of the cohorts we are going to explore how MDS-UPDRS III total is missing (whether MAR, MCAR, or MNAR)




# MDS-UPODRS III total missingness exploration - TPD

```{r}
TPD_PDonly <- P3_TPD %>% 
  filter(change_diagnosis == "")

# First, explore the missing data rates for the UPDRS variable with data in wide format.
# On a first step, we can have a look at the missingness rate for UPDRS_total
TPD_PDonly_updrstotal <- TPD_PDonly %>%
  select(ID, V1_UPDRS_III_total, V4_UPDRS_III_total, 
         V7_UPDRS_III_total, V9_UPDRS_III_total, V10_UPDRS_III_total, 
         V11_UPDRS_III_total, age_V1, age_V4, age_V7, disease_duration_diag_V1, gender)

# Looking at the missing data distribution here
summary(TPD_PDonly_updrstotal[,2:ncol(TPD_PDonly_updrstotal)])

# Exploring missing data patterns
#md.pattern(TPD_PDonly_updrstotal)

aggr_data <- aggr(TPD_PDonly_updrstotal, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, 
                  labels= names(TPD_PDonly_updrstotal), cex.axis=.5, 
                  gap=3, oma = c(10,5,5,3),
                  ylab=c("Histogram of missing data","Pattern"))

# Plostting and looking at missing data, with a special interest in combinations
#plot(aggr_data) ;
summary(aggr_data)


marginplot(TPD_PDonly_updrstotal[,c("V4_UPDRS_III_total","V7_UPDRS_III_total")])


#Looking at the distribution of missing and observed data
marginplot(TPD_PDonly_updrstotal[,c("age_V1", "V7_UPDRS_III_total")], delimiter="_imp", alpha=0.6, pch=c(19))
marginplot(TPD_PDonly_updrstotal[,c("age_V1", "V4_UPDRS_III_total")], delimiter="_imp", alpha=0.6, pch=c(19))


marginplot(TPD_PDonly_updrstotal[,c("disease_duration_diag_V1", "V7_UPDRS_III_total")], delimiter="_imp", alpha=0.6, pch=c(19), col = c("skyblue", "red"))
marginplot(TPD_PDonly_updrstotal[,c("disease_duration_diag_V1", "V4_UPDRS_III_total")], delimiter="_imp", alpha=0.6, pch=c(19), col = c("skyblue", "red"))


marginplot(TPD_PDonly_updrstotal[,c("gender","V7_UPDRS_III_total")], delimiter="_imp", alpha=0.6, pch=c(19), col = c("skyblue", "red"))
marginplot(TPD_PDonly_updrstotal[,c("gender", "V4_UPDRS_III_total")], delimiter="_imp", alpha=0.6, pch=c(19), col = c("skyblue", "red"))




```


# OPDC - MDS-UPDRS III total missingness exploration

```{r}
# First we get the UPDRSIII total available and then we transform them into long format
# to study the UDPRS total 

# Filtering df
OPDC_updrstotal <- Discovery_FU %>%
  select(subjid, visit, UPDRS_III,age_diag,
         gender) %>%
  filter(visit %in% 1:3)

# We then reshape to study
OPDC_updrstotal_wide <- OPDC_updrstotal %>%
  reshape(., idvar = "subjid", v.names = "UPDRS_III", timevar = "visit", direction= "wide") 


# Now we get the disease duration, and the MOCA variable
OPDC_updrstotal_wide = OPDC_updrstotal_wide %>%
  inner_join(Discovery_FU %>% 
               select(subjid, visit,
                      disease_duration_onset,
                      MOCA_total_adj) %>%
               filter(visit == 1) %>%
               dplyr::select(-visit), by = c("subjid"))


OPDC_updrstotal_wide <- OPDC_updrstotal_wide %>%
  select(order(colnames(OPDC_updrstotal_wide)))

# Exploring missing data patterns
#md.pattern(OPDC_updrstotal_wide)

# Work out the name axes
aggr_data <- aggr(OPDC_updrstotal_wide, col=c('navyblue','red'), numbers=TRUE,
                  sortVars=TRUE, 
                  labels= names(OPDC_updrstotal_wide), cex.axis=.5, 
                  gap=3, oma = c(10,5,5,3),
                  ylab=c("Histogram of missing data","Pattern"))



summary(aggr_data)


#Looking at the distribution of missing and observed data
marginplot(OPDC_updrstotal_wide[,c("UPDRS_III.1","UPDRS_III.2")])


#Looking at the distribution of missing and observed data
marginplot(OPDC_updrstotal_wide[,c("age_diag", "UPDRS_III.2")], delimiter="_imp", alpha=0.6, pch=c(19))
marginplot(OPDC_updrstotal_wide[,c("age_diag", "UPDRS_III.3")], delimiter="_imp", alpha=0.6, pch=c(19))


marginplot(OPDC_updrstotal_wide[,c("disease_duration_onset", "UPDRS_III.2")], delimiter="_imp", alpha=0.6, pch=c(19), col = c("skyblue", "red"))
marginplot(OPDC_updrstotal_wide[,c("disease_duration_onset", "UPDRS_III.3")], delimiter="_imp", alpha=0.6, pch=c(19), col = c("skyblue", "red"))


marginplot(OPDC_updrstotal_wide[,c("gender","UPDRS_III.2")], delimiter="_imp", alpha=0.6, pch=c(19), col = c("skyblue", "red"))
marginplot(OPDC_updrstotal_wide[,c("gender", "UPDRS_III.3")], delimiter="_imp", alpha=0.6, pch=c(19), col = c("skyblue", "red"))

marginplot(OPDC_updrstotal_wide[,c("MOCA_total_adj","UPDRS_III.2")], delimiter="_imp", alpha=0.6, pch=c(19), col = c("skyblue", "red"))
marginplot(OPDC_updrstotal_wide[,c("MOCA_total_adj", "UPDRS_III.3")], delimiter="_imp", alpha=0.6, pch=c(19), col = c("skyblue", "red"))
```


# PPMI - MDS-UPDRS III total missingness exploration

```{r}
PD_IDs <- gsub("[^0-9]", "", PPMI_PD_IDs$V1) %>% as.integer()


PPMI_Original_Cohort = PPMI_Original_Cohort %>%
  filter(PATNO %in% PD_IDs) 

PPMI_Original_Cohort_filt = PPMI_Original_Cohort %>%
  select(PATNO, EVENT_ID, updrs3_score,
         agediag, gen) 

PPMI_Original_Cohort_filt_wide = PPMI_Original_Cohort_filt %>%
   reshape(., idvar = "PATNO", v.names = "updrs3_score" , timevar = "EVENT_ID", 
          direction = "wide", sep = "_")
  

# Now we get the disease duration, and the MOCA variable
PPMI_Original_Cohort_filt_wide  = PPMI_Original_Cohort_filt_wide %>%
  inner_join(PPMI_Original_Cohort %>% 
               select(PATNO, EVENT_ID,
                      duration,
                      moca) %>%
               filter(EVENT_ID == "BL") %>%
               dplyr::select(-EVENT_ID), by = c("PATNO"))



# Exploring missing data patterns
#md.pattern(OPDC_updrstotal_wide)

# Work out the name axes
aggr_data <- aggr(PPMI_Original_Cohort_filt_wide, col=c('navyblue','red'), numbers=TRUE,
                  sortVars=TRUE, 
                  labels= names(PPMI_Original_Cohort_filt_wide), cex.axis=.5, 
                  gap=3, oma = c(10,5,5,3),
                  ylab=c("Histogram of missing data","Pattern"))

# Plostting and looking at missing data, with a special interest in combinations
#plot(aggr_data) ; 
summary(aggr_data)


# For a brief comparison, to keep consistent the missingness exploration strategy with the done for ProBanD and OPDC
# we will only compare the outcome variables at 18 and 36 months.

#Looking at the distribution of missing and observed data
marginplot(PPMI_UPDRS_PD_wide[,c("updrs3_score_V06","updrs3_score_V08")])


#Looking at the distribution of missing and observed data
marginplot(PPMI_Original_Cohort_filt_wide[,c("agediag", "updrs3_score_V06")], delimiter="_imp", alpha=0.6, pch=c(19))
marginplot(PPMI_Original_Cohort_filt_wide[,c("agediag", "updrs3_score_V08")], delimiter="_imp", alpha=0.6, pch=c(19))


marginplot(PPMI_Original_Cohort_filt_wide[,c("duration", "updrs3_score_V06")], delimiter="_imp", alpha=0.6, pch=c(19), col = c("skyblue", "red"))
marginplot(PPMI_Original_Cohort_filt_wide[,c("duration", "updrs3_score_V08")], delimiter="_imp", alpha=0.6, pch=c(19), col = c("skyblue", "red"))


marginplot(PPMI_Original_Cohort_filt_wide[,c("gender","updrs3_score_V06")], delimiter="_imp", alpha=0.6, pch=c(19), col = c("skyblue", "red"))
marginplot(PPMI_Original_Cohort_filt_wide[,c("gender", "updrs3_score_V08")], delimiter="_imp", alpha=0.6, pch=c(19), col = c("skyblue", "red"))

marginplot(PPMI_Original_Cohort_filt_wide[,c("gen","updrs3_score_V06")], delimiter="_imp", alpha=0.6, pch=c(19), col = c("skyblue", "red"))
marginplot(PPMI_Original_Cohort_filt_wide[,c("gen", "updrs3_score_V08")], delimiter="_imp", alpha=0.6, pch=c(19), col = c("skyblue", "red"))



```



# Conclussion

We have seen how, as expected, as the time increases patients withdraw from the study.
In order to determine how the data is missing, explored the MDS-UPDRS III outcome missingness against the following demographics and clinical outcome
  - Gender
  - Age at diagnosis
  - Disease duration
  - MOCA

We used VIM R pakcage to do explore missingness.
We did not find any cluster in the data missingness, suggesting data MAR.

Missing data did not cluster across any demopgrahic or MOCA we assessed in any of the cohorts

We conclude data is MCAR. Therefore, we can procede and impute missing data with the les than 20% missingness approach





