---
title: 'Imputation of splitted UPDRS measures'
author: 
- name: "Alejandro M.Carrasco"
  affiliation: UCL
output: 
  html_document:
    css: ./font/missingexploration_fontsize.css
    code_folding: hide
    theme: paper
    highlight: kate
    df_print: paged
    fontsize: 20pt
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
library(tidyverse)      # For tidy manipulation of data
library(here)           # For file path construction
library(DT)             # To display data tables
library(readxl)         # To read excel files
library(naniar)         # reaplce values with NAs on dfs
library(tableHTML)      # To generate nice html formatted table
theme_rhr <-  theme_bw(base_family = "Helvetica") + 
  theme(panel.grid.major.x = element_blank(),
        legend.position = "right",
        strip.text = element_text(size = 8),
        strip.text.y = element_text(angle = 90),
        axis.text.x = element_text(size = 8, angle = 90, hjust = 1, vjust = 0.5),
        axis.text.y = element_text(size = 8),
        axis.title.y = element_text(vjust = 0.6),
        axis.title = element_text(size = 10),
        panel.spacing = unit(0.1, "lines"))
knitr::opts_chunk$set(echo = T, warning = F, message= F)

```


# Description

Following up on the Rmarkdown number 3, we determined data to be missing completely at random.
Therefore, we decided to impute the data based on the 20% approach

Nevertheless, we noticed that one item in PD-STAT MDS-UPDRS III was MNAR. 
We took this into account in the imputation process.

# Data load

We load on out workspace all the data we are going to process

```{r}
Discovery_FU <- (read.csv("~/Data/WorkingDirectory/OPDC/clinical_data/Discovery_FU_data_with_onset_2021_04.csv"))
P3_TPD <- as_tibble(read.csv("~/Data/WorkingDirectory/PROBAND/clinical_data/P3_TPD_Version2_temp_06_16_2020.csv"))
TPD_imputed0.8 <- read.csv("../Data_processed/clinicalData_Imputed/TPD_0.8.imputed.txt")
PPMI_UPDRS<- as_tibble(read.csv("/home/amcalejandro/Data/WorkingDirectory/PPMI/clinical_data/MDS_UPDRS_Part_III.csv"))
PPMI_PD_IDs <- read.delim("/home/amcalejandro/Data/WorkingDirectory/PPMI/clinical_data/PD_IDs.txt", header=FALSE) 
PDSTAT_UPDRS <-  readxl::read_xlsx("/home/amcalejandro/Data/WorkingDirectory/PDSTAT/clinical_data/FUP_Export_CRF_MDSUPDRS_PartIII_Off.xlsx")
DIGPD <- readxl::read_xlsx("/home/amcalejandro/Data/WorkingDirectory/DIGPD_DATA/dig_pd_2019_11_Alejandro.xlsx")

PPMI_Original_Cohort <- read.csv("/home/amcalejandro/Data/WorkingDirectory/PPMI/clinical_data/PPMI_Original_Cohort_BL_to_Year_5_Dataset_Apr2020.csv")

source(here::here("R", "utils", "utils.R"))
```


### MDS UPDRS III total, limb and axial imputation


```{r}
TPD_PDonly <- P3_TPD %>% 
  filter(change_diagnosis == "")

visits_interest <- c("1","4","7")
TPD_PDonly_3visits  <- TPD_PDonly %>% 
  select(ID, gender, age_onset,
         ends_with(c(paste0("age_V", visits_interest), 
                     paste0("disease_duration_diag_V",visits_interest))),
         contains(c("time", paste0("V", visits_interest, "_", "UPDRS_III"))),
         -ends_with("total")) 

# Conveting to long format
TPD_PDonly_3visits_long <- TPD_PDonly_3visits %>% 
  gather(UPDRSIII_measure, score ,starts_with("V")) %>% 
  mutate(visit_number = str_sub(UPDRSIII_measure, 2, 2)) 


TPD_PDonly_3visits_long_updrsIII_imputed <- impute_UPDRSIII(df = TPD_PDonly_3visits_long, 
                                                           ID = "ID", 
                                                           score_column = "score", 
                                                           visit_column = "visit_number",
                                                           UPDRSIII_column = "UPDRSIII_measure",
                                                           UPDRS_type = c("total", "axial", "limb"))

write.csv(TPD_PDonly_3visits_long_updrsIII_imputed, 
          "../Data_processed/clinicalData_Imputed/TPD_0.8.imputed.txt", 
          quote = FALSE, row.names = FALSE) # 


visits_interest <- c("1","4","7", "9", "10", "11")
TPD_PDonly_AllVisits  <- TPD_PDonly %>% 
  select(ID, gender, age_onset,
         ends_with(c(paste0("age_V", visits_interest), 
                     paste0("disease_duration_diag_V",visits_interest))),
         contains(c("time", paste0("V", visits_interest, "_", "UPDRS_III"))),
         -ends_with("total")) 

# Conveting to long format
TPD_PDonly_AllVisits_long <- TPD_PDonly_AllVisits %>% 
  gather(UPDRSIII_measure, score ,starts_with("V")) %>% 
  #mutate(visit_number = str_sub(UPDRSIII_measure, 2, 2)) 
  mutate(visit_number = str_extract(UPDRSIII_measure, regex("(\\d+)")))

TPD_PDonly_AllVisits_long_updrsIII_imputed <- impute_UPDRSIII(df = TPD_PDonly_AllVisits_long, 
                                                           ID = "ID", 
                                                           score_column = "score", 
                                                           visit_column = "visit_number",
                                                           UPDRSIII_column = "UPDRSIII_measure",
                                                           UPDRS_type = c("total", "axial", "limb"))

write.csv(TPD_PDonly_AllVisits_long_updrsIII_imputed, 
          "../Data_processed/clinicalData_Imputed/TPD_0.8.imputed.AllVisits.txt", 
          quote = FALSE, row.names = FALSE)

```


### OPDC imputation


```{r}
# Imputing the values
Discovery_FU_3visits <- Discovery_FU %>%
  select(subjid, visit, gender, age_onset, age_diag, age, UPDRS_III,
         visit, starts_with("af_3"), -contains(c("3_a","3_b", "3_c"))) %>%
  filter(visit %in% 1:3)

# Converting to long format
Discovery_FU_3visits_long <- Discovery_FU_3visits %>% 
  gather(UPDRSIII_measure, score ,starts_with("af_")) 

OPDC_3visits_long_updrsIII_imputed <- impute_UPDRSIII(df = Discovery_FU_3visits_long, 
                                                           ID = "subjid", 
                                                           score_column = "score", 
                                                           visit_column = "visit",
                                                           UPDRSIII_column = "UPDRSIII_measure",
                                                           UPDRS_type = c("total", "axial", "limb"))

write.csv(OPDC_3visits_long_updrsIII_imputed, 
          "../Data_processed/clinicalData_Imputed/OPDC_0.8.imputed.txt", 
          quote = FALSE, row.names = FALSE)


# We get all the scores up to 90 months for a further exploration of the rate of progression
Discovery_FU_all <- Discovery_FU %>%
  select(subjid, visit, gender, age_onset, age_diag, age, UPDRS_III,
         visit, starts_with("af_3"), -contains(c("3_a","3_b", "3_c"))) %>%
  filter(visit %in% 1:7)

# Converting to long format
Discovery_FU_all_long <- Discovery_FU_all %>% 
  gather(UPDRSIII_measure, score ,starts_with("af_")) 

OPDC_all_long_updrsIII_imputed <- impute_UPDRSIII(df = Discovery_FU_all_long, 
                                                           ID = "subjid", 
                                                           score_column = "score", 
                                                           visit_column = "visit",
                                                           UPDRSIII_column = "UPDRSIII_measure",
                                                           UPDRS_type = c("total", "axial", "limb"))

write.csv(OPDC_all_long_updrsIII_imputed,
          "../Data_processed/clinicalData_Imputed/OPDC_0.8.imputed.AllVisits.txt", 
          quote = FALSE, row.names = FALSE)

```



### PPMI imputation

```{r}
# Keeping PD only and renaming a misspelled variable
PD_IDs <- gsub("[^0-9]", "", PPMI_PD_IDs$V1) %>% as.integer()
PPMI_UPDRS_PD <- subset(PPMI_UPDRS, PATNO %in% PD_IDs)
colnames(PPMI_UPDRS_PD)[17] <- "NP3RIGRL"


PPMI_UPDRS_PD_3 <- PPMI_UPDRS_PD %>% 
  filter((EVENT_ID %in% c("BL", paste0("V0",1:8)))) %>%
  gather(UPDRSIII_measure, score ,starts_with("NP3"))  

PPMI_long_updrsIII_imputed <- impute_UPDRSIII(df = PPMI_UPDRS_PD_3,
                                                           ID = "PATNO", 
                                                           score_column = "score", 
                                                           visit_column = "EVENT_ID",
                                                           UPDRSIII_column = "UPDRSIII_measure",
                                                           UPDRS_type = c("total", "axial", "limb"))


PPMI_long_updrsIII_imputed <- PPMI_long_updrsIII_imputed %>% 
  mutate(UPDRSIII_measure_total = str_remove(UPDRSIII_measure_total, "."),
         UPDRSIII_measure_axial = str_remove(UPDRSIII_measure_axial, "."),
         UPDRSIII_measure_limb = str_remove(UPDRSIII_measure_limb, "."))

write.csv(PPMI_long_updrsIII_imputed, 
          "../Data_processed/clinicalData_Imputed/PPMI_0.8.imputed.txt", 
          quote = FALSE, row.names = FALSE)



# Now we get data for all the visits
PPMI_UPDRS_PD_all<- PPMI_UPDRS_PD %>%
  filter(!EVENT_ID %in% c("SC", "ST", "PW", "U01")) %>%
  gather(UPDRSIII_measure, score ,starts_with("NP3")) 


PPMI_long_updrsIII_all_imputed <- impute_UPDRSIII(df = PPMI_UPDRS_PD_all_age,
                                                           ID = "PATNO", 
                                                           score_column = "score", 
                                                           visit_column = "EVENT_ID",
                                                           UPDRSIII_column = "UPDRSIII_measure",
                                                           UPDRS_type = c("total", "axial", "limb"))

PPMI_long_updrsIII_all_imputed <- PPMI_long_updrsIII_all_imputed %>% 
  mutate(UPDRSIII_measure_total = str_remove(UPDRSIII_measure_total, "."),
         UPDRSIII_measure_axial = str_remove(UPDRSIII_measure_axial, "."),
         UPDRSIII_measure_limb = str_remove(UPDRSIII_measure_limb, "."))

write.csv(PPMI_long_updrsIII_all_imputed, "../Data_processed/clinicalData_Imputed/PPMI_0.8.imputed.AllVisits.txt", quote = FALSE, row.names = FALSE) # 
```


## PD-STAT imputation

For PDSTAT we have UPDRS III measures spaced 12 months instead of 18 months as for the previous cohorts.
Furthermore, we only have data up to 24 months from BL. We use the three data points available ( BL, 12m, 24m)

Worth noting that : "In PD STAT, one item of the MDS-UPDRS part III was erroneously omitted from case report forms at all time-points. Question 15 should consist of two items, a score for postural tremorfor each of the left and right hands. However, only one item response could be recorded and the case report form did not specify, or request specification of, which hand. For a small proportion of participants, the assessor manually annotated the case report forms and included postural tremor scores for the left and right hand. In these instances, the scores for both left and right hands will be used as part of the calculation of the primary outcome. However, if a participant only has one postural tremor score, the second postural tremor score will be treated as missing data (not at random). "


```{r}
source(here::here("R", "imputation_UPDRS.R"))

PDSTAT_UPDRS_3visits_tremorHands  <- PDSTAT_UPDRS %>% 
  select(StudyNumber, starts_with("M"), 
         -ends_with(c("Total","PNC",
                      "Missing","DyskinesiasPresent",
                      "DyskinesiaMovementsIntefere", "HAndYStage"))) %>%
  mutate(
    M00_MDSUPDRS_PartIII_Off_PosturalTremorHandsRight = case_when(
      ((M00_MDSUPDRS_PartIII_Off_PosturalTremorHandsRight == 999) & 
         (M00_MDSUPDRS_PartIII_Off_PosturalTremorHandsLeft == 999)) ~ M00_MDSUPDRS_PartIII_Off_PosturalTremorHands,
     ((M00_MDSUPDRS_PartIII_Off_PosturalTremorHandsRight != 999) & 
         (M00_MDSUPDRS_PartIII_Off_PosturalTremorHandsLeft != 999)) ~ M00_MDSUPDRS_PartIII_Off_PosturalTremorHandsRight),
    
    M12_MDSUPDRS_PartIII_Off_PosturalTremorHandsRight = case_when(
      ((M12_MDSUPDRS_PartIII_Off_PosturalTremorHandsRight == 999) & 
       (M12_MDSUPDRS_PartIII_Off_PosturalTremorHandsLeft == 999)) ~ M12_MDSUPDRS_PartIII_Off_PosturalTremorHands,
       ((M12_MDSUPDRS_PartIII_Off_PosturalTremorHandsRight != 999) & 
         (M12_MDSUPDRS_PartIII_Off_PosturalTremorHandsLeft != 999)) ~ M12_MDSUPDRS_PartIII_Off_PosturalTremorHandsRight),
    
    M24_MDSUPDRS_PartIII_Off_PosturalTremorHandsRight = case_when(
      ((M24_MDSUPDRS_PartIII_Off_PosturalTremorHandsRight == 999) & 
       (M24_MDSUPDRS_PartIII_Off_PosturalTremorHandsLeft == 999)) ~ M24_MDSUPDRS_PartIII_Off_PosturalTremorHands,
      ((M24_MDSUPDRS_PartIII_Off_PosturalTremorHandsRight != 999) & 
         (M24_MDSUPDRS_PartIII_Off_PosturalTremorHandsLeft != 999)) ~ M24_MDSUPDRS_PartIII_Off_PosturalTremorHandsRight)
    ) %>%
  select(-ends_with("PosturalTremorHands"))


# Getting the df to look at the proportion of missingness before imputation
PDSTAT_UPDRS_3visits_UPDRStotal  <- PDSTAT_UPDRS %>% 
  select(StudyNumber, ends_with("Total"))


# Check we only have 33 scores
PDSTAT_UPDRS_3visits_tremorHands %>% select(starts_with("M00")) %>% length()
  
PDSTAT_UPDRS_3visits_tremorHands_long <- PDSTAT_UPDRS_3visits_tremorHands %>% 
  gather(UPDRSIII_measure, score ,starts_with("M")) %>% 
  mutate(visit_number = str_sub(UPDRSIII_measure, 2, 3),
         UPDRSIII_measure = gsub("^.*_", "", UPDRSIII_measure)) %>%
  naniar::replace_with_na_all(condition = ~.x == 999) # We set 999 as NA


# In our case, one score for postural tremor of the hands is consistently missing
# We use this score for limb and total outcome. Therefore, we will have to adjust the
# missingness allowance according to having one variable consistently missing.


PDSTAT_3visits_imputed_newThreshold <- impute_UPDRSIII(df = PDSTAT_UPDRS_3visits_tremorHands_long, 
                                                           ID = "StudyNumber", 
                                                           score_column = "score", 
                                                           visit_column = "visit_number",
                                                           UPDRSIII_column = "UPDRSIII_measure",
                                                           UPDRS_type = c("total", "axial", "limb"),
                                                           missing_threshold = c(3,2,1))

write.csv(PDSTAT_3visits_imputed_newThreshold, "../Data_processed/clinicalData_Imputed/PDSTAT_0.8.imputed.txt",  
          quote = FALSE, row.names = FALSE) 

```


## DIGPD imputation


```{r}
# Attaching the function
source(here::here("R", "imputation_UPDRS.R"))

DIGPD_3visits <- DIGPD %>% filter(visite %in% 1:4) %>% 
  dplyr::mutate(across(starts_with("MOTOREX"), ~ as.numeric(.x)))

# Getting data needed in wide format for further missingnesss exploration
DIGPD_wide <- DIGPD_3visits %>% 
  select(PATIENT, visite, SCOREIII) %>% mutate(visite = as.character(visite)) %>% as.data.frame() %>%
  stats::reshape(., idvar = "PATIENT", v.names = "SCOREIII", timevar = "visite", 
                 varying = list(c("SCOREIII_1", "SCOREIII_2", "SCOREIII_3", "SCOREIII_4")), direction= "wide") 

# Converting to long format
DIGPD_3visits_long <- DIGPD_3visits %>% 
  gather(UPDRSIII_measure, score ,starts_with("MOTOREXAM")) 

DIGPD_3visits_long_updrsIII_imputed <- impute_UPDRSIII(df = DIGPD_3visits_long, 
                                                           ID = "PATIENT", 
                                                           score_column = "score", 
                                                           visit_column = "visite",
                                                           UPDRSIII_column = "UPDRSIII_measure",
                                                           UPDRS_type = c("total", "axial", "limb"))



# Comment out if not existing
write.csv(DIGPD_3visits_long_updrsIII_imputed, "../Data_processed/clinicalData_Imputed/DIGPD_0.8.imputed.txt", quote = FALSE, row.names = FALSE)


DIGPD_all <- DIGPD %>% 
  dplyr::mutate(across(starts_with("MOTOREX"), ~ as.numeric(.x)))

# Getting data needed in wide format for further missingnesss exploration
DIGPD_wide_all <- DIGPD_all %>% 
  select(PATIENT, visite, SCOREIII) %>% mutate(visite = as.character(visite)) %>% as.data.frame() %>%
  stats::reshape(., idvar = "PATIENT", v.names = "SCOREIII", timevar = "visite", 
                 varying = list(c("SCOREIII_1", "SCOREIII_2", "SCOREIII_3", "SCOREIII_4",
                                  "SCOREIII_5","SCOREIII_6", "SCOREIII_7")), direction= "wide") 

# Converting to long format
DIGPD_all_long <- DIGPD_all %>% 
  gather(UPDRSIII_measure, score ,starts_with("MOTOREXAM")) 

DIGPD_all_long_updrsIII_imputed <- impute_UPDRSIII(df = DIGPD_all_long, 
                                                           ID = "PATIENT", 
                                                           score_column = "score", 
                                                           visit_column = "visite",
                                                           UPDRSIII_column = "UPDRSIII_measure",
                                                           UPDRS_type = c("total", "axial", "limb"))

write.csv(DIGPD_all_long_updrsIII_imputed, "../Data_processed/clinicalData_Imputed/DIGPD_0.8.imputed.AllVisits.txt", quote = FALSE, row.names = FALSE)
```



# Generate tables to explore the missingness rates before and after imputation

```{r}

# Missingness after imputation
# ProBanD
TPD_missprop_afterIMputation <- TPD_PDonly_3visits_long_updrsIII_imputed %>% 
  group_by(visit_number) %>%
  dplyr::summarise(across(c("UPDRSIIItotal_imputed"), ~ (sum(is.na(.x)) + (1963 - n())) / 1963))
# OPDC
OPDC_missprop_afterIMputation <- OPDC_3visits_long_updrsIII_imputed %>% 
  group_by(visit) %>%
  dplyr::summarise(across(c("UPDRSIIItotal_imputed"), ~ (sum(is.na(.x)) + (963 - n())) / 963))
# PPMI
PPMI_missprop_afterIMputation <- PPMI_long_updrsIII_imputed %>% 
  group_by(EVENT_ID) %>%
  dplyr::summarise(across(c("UPDRSIIItotal_imputed"), ~ (sum(is.na(.x)) + (415 - n())) / 415))
#PDSTAT
PDSTAT_missprop_afterImputation <- PDSTAT_3visits_tremoshands_long_updrsIII_imputed_newThreshold %>%
  group_by(visit_number) %>%
  dplyr::summarise(across(c("UPDRSIIItotal_imputed"), ~ (sum(is.na(.x)) + (174 - n())) / 174))
#DIGPD
DIGPD_missprop_afterImputation <- DIGPD_3visits_long_updrsIII_imputed  %>%
  group_by(visite) %>%
  dplyr::summarise(across(c("UPDRSIIItotal_imputed"), ~ (sum(is.na(.x)) + (415 - n())) / 415))


# Cohorts with 18 months spaced data available
imputation_summary = data.frame(t(data.frame("PROBAND" = sapply(TPD_PDonly_updrstotal, function(x) sum(is.na(x)) / length(x))[2:4],
                                  "OPDC" = sapply(OPDC_updrstotal_wide, function(x) sum(is.na(x)) / length(x))[2:4],
                                  "PPMI" = sapply(PPMI_UPDRS_PD_wide[, c(2,7,10)], function(x) sum(is.na(x)) / length(x)),
                                  "PROBAND" = TPD_missprop_afterIMputation$UPDRSIIItotal_imputed,
                                  "OPDC" = OPDC_missprop_afterIMputation$UPDRSIIItotal_imputed,
                                  "PPMI" = PPMI_missprop_afterIMputation$UPDRSIIItotal_imputed[c(1,6,8)],
                                  row.names = c("BL","18 months","36 months")))) %>% 
  rownames_to_column(var = "Cohort") %>% as_tibble() %>%
  mutate_if(is.numeric, round, digits=4) 
# Some wrangling
imputation_summary$Cohort <- str_remove(imputation_summary$Cohort, "[\\.].*")

tableHTML(imputation_summary, 
          rownames = FALSE, 
          row_groups = list(c(3, 3), c('Before imputation', "After imputation")),
          second_headers = list(c(1, 4), c('Imputation status', 'missing rate at different time points')),
          collapse = 'separate') #%>% tableHTML_to_image(file = "/home/alejandromc/Kronos_downloads/prueba",type = "jpeg")


# Cohorts with 12 months spaced data available
imputation_summary_2 = data.frame(t(data.frame("PDSTAT" =
                                                 c(sapply(PDSTAT_UPDRS_3visits_UPDRStotal, function(x) sum(is.na(x)) / length(x))[2:4],NA),
                                               "DIGPD" = sapply(DIGPD_wide, function(x) sum(is.na(x)) / length(x))[2:5],
                                               "PDSTAT" = c(PDSTAT_missprop_afterImputation$UPDRSIIItotal_imputed, NA),
                                               "DIGPD" = DIGPD_missprop_afterImputation$UPDRSIIItotal_imputed,
                                               row.names = c("BL","12 months","24 months", "36 months")))) %>%
  rownames_to_column(var = "Cohort") %>% as_tibble() %>%
  mutate_if(is.numeric, round, digits=4) 

imputation_summary_2$Cohort <- str_remove(imputation_summary_2$Cohort, "[\\.].*")  

tableHTML(imputation_summary_2, 
          rownames = FALSE, 
          row_groups = list(c(2, 2), c('Before imputation', "After imputation")),
          second_headers = list(c(1, 5), c('Imputation status', 'missing rate at different time points')),
          collapse = 'separate') #%>% tableHTML_to_image(file = "/home/alejandromc/Kronos_downloads/prueba",type = "jpeg")

```
